---
title: "Multi session analysis 2019"
author: "Fernando Mateos-Gonzalez (ALKA Wildlife)"
output:
  word_document: default
  html_document: default
  html_notebook: default
  pdf_document: default
---

This report analises data from capture-recapture sessions during 17-18-19 April, 24-25 April and 15-19, 24 July 2019, in order to calculate the density of the population.




```{r results="hide"}
library(secr)
library(here)

```

## Analysis

This analysis was carried out using the package 'secr' version 3.2.1 and  R `r getRversion()`

We start by creating the capthist, the file combining our captures with the trap locations:



```{r}

multi <- read.capthist(here("data", "multicaptures.txt"), here("data", "traps.txt"), detector = "proximity")

summary(multi)


```

*n number of distinct individuals detected on each occasion t  
*u number of individuals detected for the first time on each occasion t  
*f number of individuals detected on exactly t occasions  
*M(t+1) cumulative number of detected individuals on each occasion t  



Now we use the plot method, which for capthist objects has additional arguments; we set tracks
= TRUE to join consecutive captures of each individual.



```{r}
par(mfrow = c(1,1), mar = c(1,1,1,1)) # reduce margins
plot (multi, tracks = TRUE, gridsp = 5,border = 10)
      

```




The most important insight from this figure is that individuals tend to be recaptured near their site of first capture. This is expected when the individuals of a species occupy home ranges. In SECR models the
tendency for detections to be localised is reflected in the spatial scale parameter σ. 

Successive trap-revealed movements can be extracted with the moves function and summarised with hist:


```{r}
m <- unlist(moves(multi))
par(mar = c(3.2,4,1,1), mgp = c(2.1,0.6,0)) # reduce margins
hist(m, breaks = seq(0/5, 50,5), xlab = "Movement m", main = "")
```

We will employ the estimate of the spatial scale σ from the July session (6.21) to fit the simplest possible SECR model with function secr.fit. 


```{r}
fit <- secr.fit (multi, buffer = 4 * 6.21, trace = FALSE,biasLimit = NA, verify = FALSE)

detector(traps(multi)) <- "proximity"

fit
```


The report comprises:  
• function call and time stamp  
• summary of the data  
• description of the model, including the maximized log likelihood, Akaike’s Information Criterion AIC  
• estimates of model coefficients (beta parameters)  
• estimates of variance-covariance matrix of the coefficients  
• estimates of the ‘real’ parameters  

. The estimated density is 266 susliks per hectare, 95% confidence interval 196-360 susliks per hectare  
. The other two real parameters jointly determine the detection function, plotted below with 95% confidence limits



```{r}
par(mar = c(4,4,1,1)) # reduce margins
plot(fit, limits = TRUE)


```




The theory of SECR tells us that buffer width is not critical as long as it is wide enough that animals at
the edge have effectively zero chance of appearing in our sample.We check that for the present model with the function esa.plot.  
The estimated density has easily reached a plateau at the chosen buffer width (dashed red line):



```{r}
esa.plot(fit)
abline(v = 4 * 6.21, lty = 2, col = 'red')

```

dbar is the mean distance between consecutive capture locations, pooled over individuals (e.g. Efford 2004). moves returns the raw distances.

MMDM (for ‘Mean Maximum Distance Moved’) is the average maximum distance between detections of each individual i.e. the observed range length averaged over individuals (Otis et al. 1978).

ARL (or ‘Asymptotic Range Length’) is obtained by fitting an exponential curve to the scatter of observed individual range length vs the number of detections of each individual (Jett and Nichols 1987: 889).

RPSV (for ‘Root Pooled Spatial Variance’) is a measure of the 2-D dispersion of the locations at which individual animals are detected, pooled over individuals (cf Calhoun and Casby 1958, Slade and Swihart 1983).

moves reports the distance between successive detections of each animal.

```{r}

dbar(multi, userdist = NULL, mask = NULL)
MMDM(multi, min.recapt = 1, full = FALSE, userdist = NULL, mask = NULL)
RPSV(multi, CC = FALSE)
moves(multi, userdist = NULL, mask = NULL, names = FALSE)
```


